Implementation of a stable, high-power optical lattice for quantum gas microscopy

### 2018-06-13 LatticeOptics/latticeOpticsv5/LatticeOpticsV5.tex
Many changes to abstract, text, fixes to figures

\newcommand\unit[2]{\ensuremath{#1~\mathrm{{#2}}}}

\newcommand\Ket[1]{\ensuremath{|{#1}\rangle}}
\newcommand\Bra[1]{\ensuremath{\langle{#1}|}}

\newcommand\Isotope[2]{\ensuremath{^{#1}\mathrm{#2}}}
\newcommand\Li{\Isotope{6}{Li}}

\renewcommand{\vec}[1]{\ensuremath{\bm{#1}}}

\newcommand\NA{\ensuremath{\mathrm{NA}}}
\newcommand\FIXME{{\color{red}\ensuremath{\mathrm{FIXME}}}}

@@

\author{A. Mazurenko}
\altaffiliation{
  Present address: FIXME}
\author{S. Blatt}
\altaffiliation{
  Present address: Max-Planck-Institut f{\"u}r Quantenoptik,
  Hans-Kopfermann-Stra{\ss}e 1, 85748 Garching, Germany}
\author{F. Huber}
\altaffiliation{
  Present address: IPG Photonics Corporation, Oxford, MA 01540}
\author{M. F. Parsons}
\altaffiliation{
  Present address: FIXME TOO}
\author{C. S. Chiu}
\author{G. Ji}
\author{D. Greif}
\author{M. Greiner}
\affiliation{
  Department of Physics, Harvard University,
  Cambridge, Massachusetts, 02138, USA}

\date{\today}
\begin{abstract}
 We describe the design of a stable high power \unit{1064}{nm} laser system for practical use in optical lattice experiments. The system is based on a low noise laser amplified by an array of four heavily modified, high-power fiber amplifiers. The beam intensity is stabilized with a nonlinear feedback loop. The position of the produced optical lattice is monitored in real time and its stability is characterized.
\end{abstract}
\maketitle

Optical lattices holding an array of ultracold atoms have become a powerful experimental platform for a wide variety of investigations into various forms of quantum simulation~\cite{Friedenauer2008, Kim2010, Struck2011, Simon2011, Yan2013, Drewes2016, Murmann2015}.
The recent advent of quantum gas microscopy~\cite{Bakr2009, Sherson2010, Haller2015, Cheuk2015, Parsons2015, Edge2015, Omran2015, Greif2016, Cheuk2016, Parsons2016, Boll2016, Cheuk2016a, Brown2017} has enabled studies that exhibit unprecedented control over individual atoms at extremely low energy scales, particularly those governed by the Hubbard Hamiltonian~\cite{Hubbard1963}.
However, this progress has placed ever more stringent technical constraints on the lasers used to trap and manipulate the atomic systems.
For example, heating due to the intensity noise cannot be allowed to approach the longest thermalization timescales, lest the system fail to thermalize~\cite{Savard1997}.
Even if the system thermalizes, heating rates can limit the interrogation time available to probe a low-temperature state.
Similarly, experiments that rely on a small number of atoms held by multiple traps place stringent requirements on the relative positional stability of the traps, to ensure that the operations performed by the traps have adequate fidelity.

In this work we discuss major improvements to our high power optical lattice system~\cite{Blatt2015} and show levels of intensity noise below \unit{-120}{dBc/Hz} between \unit{1}{kHz} and \unit{3}{MHz} Fourier frequency.
The stability of the lattice position is below the single site level (over a few hours) and the stability of the harmonic confinement is at the few site level (over nine months).

@@

The architecture of the system is shown in Fig.~\ref{fig:high_level} and is based on a single low-power, low-noise seed laser that supplies light to an array of four high-power fiber amplifiers (one for a transport beam~\cite{Huber2014}, and one each for the $X,Y,Z$ lattices~\cite{Blatt2015,Parsons2016}).
These amplifiers generate enough light to perform site resolved imaging~\cite{Parsons2015}, but also substantially increase the level of intensity and pointing noise.
The fiber amplifier output is controlled by two feedback loops, which stabilize the light in low power and high power regimes, respectively, as discussed later.
After intensity stabilization, the laser beam is mode-shaped to a desired waist.
The optical system is engineered to be passively stable in terms of pointing, but the position and phase of the optical lattice are monitored in real time, and beam pointing can be controlled with remotely actuated mirrors.

@@

\section{Laser system}
\subsection{Seed laser system}
The lattice laser system is seeded by an Innolight Mephisto laser, which provides approximately \unit{2}{W} of \unit{1064}{nm} light.
This commercial laser utilizes a non-planar ring oscillator (NPRO) design that ensures superb phase and intensity stability.
This source is split and coupled into four single-mode, polarization-maintaining fibers (Thorlabs PM980-XP), each connected to the input of a fiber amplifier.
Three of these fiber amplifiers are used to produce the 3D optical lattice~\cite{Blatt2015,Parsons2016}, and a fourth is used for transport~\cite{Huber2014}.
To avoid interference of the lattice beams, two of the beam paths pass through acousto-optic modulators (AOMs), which detune the lattice frequencies from each other.

\begin{figure*}
  \begin{center}
    \includegraphics{Figure2.pdf}
    \caption{\textbf{Laser noise of the lattice system.} (a) Shows the laser noise of unmodified Nufern fiber amplifiers and the Mephisto seed, as well as an independent Mephisto MOPA 55W (not used for the optical lattice presented in this manuscript). (b) Shows the open (OL) and closed (CL) loop behavior of the high power feedback loop. (c) Shows open loop behavior of the low power feedback loop after making the modifications. (d) Closed loop behavior of the loop system without additional filtering, measured on the same setup as panel (c).}\label{fig:noises}
  \end{center}
\end{figure*}

\subsection{Fiber amplifier system}
Quantum gas microscopy of \Li{} requires single site trap frequencies of approximately \unit{1}{MHz}~\cite{Parsons2015} to be in the Lamb-Dicke regime necessary for Raman sideband cooling.
Such trap frequencies may only be attained using tightly focused, high-power beams.
In practical terms, this requires approximately $30\,$W per lattice axis, which we attain using Nufern fiber amplifiers (NuAmp SUB1174-34).
These amplifiers can amplify $150\,$mW of seed light to up to approximately $45\,$W. They use a length of doped fiber pumped with $808\,$nm light from fiber coupled diode sources to amplify $1064\,$nm light.
The amplification is performed using a two stage design, where the stages are controlled and powered using separate power supplies and electronics.

\begin{figure*}[t]
  \begin{center}
    \includegraphics{Figure3.pdf}
    \caption{\textbf{Optical layout} of the optical lattice system for beam shaping and modulation.}
    \label{fig:optical_layout}
  \end{center}
\end{figure*}

@@

Further sources of technical noise arise from the design of the amplifier. There are two power supplies in the fiber amplifier: (1) a general purpose power supply (PSU) for the control electronics and the first, low-power amplification stage and (2) a high-power supply for the second, high-power stage.
Despite the presence of a water-cooled cold-plate for the gain fiber and pump diodes, the high power PSU is air cooled by an internal fan with variable frequency.
This fan adds noise at acoustic frequencies and degrades pointing stability by increasing vibrations on the output fiber tip.
To reduce acoustic coupling and reduce switching power supply noise, both power supplies were removed from the fiber amplifier chassis, and the low-power supply was replaced by a linear PSU (Acopian A24MT350M).
Further, a high current low-pass filter (MPE DS26387) was added to the high-power PSU (Lumina LDD-600-60-10-5VP/M).

These upgrades serve to suppress many of the noise spurs measured in the fiber amplifier RIN.
The effects can be as seen by comparing Fig.~\ref{fig:noises} (a) and (b), particularly in the region of approximately \unit{10}{kHz} Fourier frequency.
The measurements in panels (a)-(d) were carried out on a single fiber amplifier, prior to and following the modifications. The initial and final noise spectra were verified to be consistent between amplifiers coming from the same manufacturing batch.
The figure is based on fiber amplifiers from a relatively new batch, which, even in its unmodified configuration does not exhibit a large switching supply spur.
This is not the case for certain older batches, where the spur is approximately \unit{30}{dB} above the background.

The SBS noise is fundamental to the fiber amplifier architecture and cannot be entirely avoided. However, since it is a function of fiber length and optical power, it can be minimized. To this end, we use custom-ordered amplifiers with a \unit{50}{cm} long output fiber, compared to the standard length of a few meters.
In addition, we use the lowest pump diode current that yields enough power to perform imaging.
The rise of the noise floor as a function of the pump diode current can be seen in Fig.~\ref{fig:noises}(a) at Fourier frequencies above \unit{100}{kHz}.

The operation of the Nufern fiber amplifiers can be further optimized by finely controlling the cooling water.
The water temperature tunes the wavelength of the light supplied by the diodes that optically pump the gain fiber.
Optimizing the temperature is advantageous because (1) the maximum output power of the fiber amplifier is increased due to the improved efficiency of pump light conversion and (2) the lifetime of the amplifier is increased because less stray pump light needs to be absorbed when the pump light is separated from the output, reducing the thermal load.

@@

The large-diameter fiber tip at the output of the fiber amplifier is mounted in a monolithic mount made from copper.
The beam is collimated using an $f=\unit{20}{mm}$ fused-silica, air-spaced triplet collimator (Opto-Sigma HFTLSQ-15-20PF1) angled with respect to and centered on the angle-cleaved fiber tip.
The polarization is set using an IBS coated thin-film plate polarizer (Precision photonics PO1056-FY), which has the added benefit of rejecting the undesired cladding modes of the fiber, which carry about \unit{1}{W} of light.
The fact that the lattice is retro-reflected, combined with the observation that optical isolation falls with applied optical power~\cite{Yoshida1999} means that two stages of optical isolators must be used.
The optical isolators are based on \unit{5}{mm} aperture terbium gallium garnet (TGG) isolators (Thorlabs IO-5-1064-VHP), where we replaced the default polarizers (BK7 Brewster plates) by IBS-coated fused silica plate polarizers.
The isolators are located approximately \unit{1}{m} away from the ultracold atoms which are susceptible to stray magnetic fields.
To overcome this problem the isolators are enclosed in mu-metal shielding.
Actuation of the power is performed after the optical isolators, such that the full fiber amplifier output continuously passes through them, minimizing thermal transients during the experimental sequence.

@@

After the beam has been cleaned up and isolated, it is important to consider two aspects of the desired experiment:
\begin{enumerate}
\item The lattice power must be continuously tunable from around \unit{10}{\mu{}W} to around \unit{20}{W}, corresponding to lattice depths in the approximate range of $10^-3$ to $2 \times 10^3 E_R$, where $E_R$ is the geometric recoil energy corresponding to the lattice spacing~\cite{Blatt2015}.
\item The location of the minimum Gaussian beam waist must be accurately positioned to overlap the other lattices and dipole traps present in the experiment, and must remain aligned for the duration of the data collection process (hours to days).
\end{enumerate}
To satisfy the first of these requirements, a two-stage feedback system is implemented.
Two stages are used because the experiment typically operates in one of two regimes, which we term \textit{detection} and \textit{interaction}.
In the interaction phase of the experiment, the lattice is relatively shallow ($\approx \unit{100}{mW}$, corresponding to depths of $\approx 10 E_R$), which allows atoms to tunnel and interact with each other.
In this regime, precise and potentially fast control is required.
In the detection phase, the depth of the lattice is dramatically raised (to $\approx 2000 E_R$), which isolates atoms in their individual wells so that Raman sideband imaging can be performed~\cite{Parsons2015}.
In the detection regime, the control need not be fast, and the passive stability of the Mephisto laser ensures that noise is low at relevant frequency scales ($\approx\unit{2}{MHz}$ Fourier frequency).
Thus, two control loops are used, a fast loop based on an acousto-optic modulator that controls the laser at low powers, and a slow loop which uses a Berek compensator to control the laser at high powers.

After the isolation stage, we use the sequence of optical elements shown in Fig.~\ref{fig:optical_layout}, consisting of a polarizer, half-wave plate (HWP), Berek compensator mounted on a precision galvo (Thorlabs GVS002, or Camtech 8320K, depending on the beam path) half-wave plate, and polarizer.
The Berek compensator is a $z$-cut quartz plate anti-reflection-coated at \unit{1064}{nm}.
By tilting the plate about its vertical axis by a few degrees, the extraordinary axis is mixed into the propagation of the beam, leading to  tilt-dependent birefringence that can be tuned from a zero-wave plate past a half-wave plate, represented by a retardation angle $\phi$.
Combined with the subsequent polarization optics, rotation of the Berek compensator can vary the transmitted power with a contrast of $\sim$1000:1.
Figure~\ref{fig:berek} shows the transmitted power as a function of retardation angle $\phi$ and fast-axis angle $\theta$ of the waveplates surrounding the compensator.
\begin{figure}
  \begin{center}
    \includegraphics{Figure4.pdf}
    \caption{\textbf{Berek compensator} transmission as a function of the Berek compensator's retardation angle $\phi$ and waveplate rotation angles $\theta$.}
    \label{fig:berek}
  \end{center}
\end{figure}

@@

\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{Figure5.pdf}
    \caption{\textbf{Berek compensator based feedback loop}. This Berek compensator based feedback loop is used to vary the beam intensity at relatively high power. (a) Shows the small signal impulse response and (b-c) show the corresponding Bode plots of the transfer function $G$.}\label{fig:berek_step_response}
  \end{center}
\end{figure}
A key point, as seen in Fig.~\ref{fig:berek}, is that if the waveplates are rotated together, the maximum transmitted power is fixed, while the minimum can be raised.
The waveplates are detuned such that the maximum power needed in \textit{interaction} mode is the minimum power transmitted by the Berek compensator system, strongly suppressing fluctuations due to angle positioning noise of the Berek compensator.
Naturally, rotation of a parallel plate in the beam path induces a shift of the beam, but this is acceptable because the Berek compensator is only active during the imaging phase of the experiment, when the position of the underlying harmonic trap is largely irrelevant.
Fortunately, the spatial phase of the optical lattice is set by the retro-reflector mirror, which does not move, and is not affected by the rotation of the Berek compensator.
This system functions as the actuator arm of a digital feedback loop, where the power of the beam is measured using a low noise photodiode. This feedback loop regulates the lattice between $1$ W of optical power and its maximum value (which depends on the amplifier, but is $\approx20$ W).
The step response and corresponding Bode plots~\cite{Bechhoefer2005} are shown in Fig.~\ref{fig:berek_step_response}.
A polarization rotation by $2 \pi$ is produced by approximately $3^\circ$ of axial rotation of the quartz plate.
Since we never need to rotate the plate outside this range, the position is constrained to this window in software.
Since two feedback loops are present in the system (the second one detailed in the next section), care must be taken so that they do not function simultaneously.
To accomplish this, the low power loop is always manually railed to its high extreme when the high power loop is active, and the digital high power loop is disabled when the low power loop is active.

\subsection{Low power feedback}
The high power feedback used in the \textit{imaging} phase of the experiment is exceedingly robust because the heating rate from scattered resonant imaging light far exceeds any technical heating from the loop.
Even if the loop was exceedingly noisy, the heating would be entirely compensated by the simultaneous Raman cooling~\cite{Parsons2015}.

In contrast, the low power loop that controls the lattice laser during the \textit{interaction} phase of the experiment is critically important:
\begin{enumerate}
  \item Its RIN can limit the system temperature.
  \item It can affect the pointing.
  \item It must have closed loop bandwidths of $\geq1\mathrm{\,kHz}$, and open loop modulation bandwidths $\geq 100\mathrm{\,kHz}$.
\end{enumerate}
To this end, we use an all-custom loop based on a TeO$_2$ AOM (Gooch \& Housego AOMO 3080-198), where the power of the RF supplied to the AOM is used to actuate the fast, low power feedback loop.
The AOM was chosen for its low thermal lensing and easily-accessible RF powers.
It is mounted on a custom-built, monolithic, water-cooled flexure mount that allows optimization of the AOM efficiency, that has been found to remain stable over years of operation.

\begin{figure}
  \begin{center}
    \includegraphics{Figure6.pdf}
    \caption{\textbf{Feedback circuits.} (a) The reverse-biased photodiode. (b) Externally referenced phase-locked loop. (c) Nonlinear loop filter. (d) RF power modulation using a mixer.}\label{fig:circuits}
  \end{center}
\end{figure}
The feedback circuit is shown in Fig.~\ref{fig:circuits}a. As seen in Fig.~\ref{fig:optical_layout}, a pair of low noise photodiodes measures the optical power of the beam sampled with a weakly reflective optic (Thorlabs BSF10-C).
The beam sampler is wedged in such a way that the two reflections diverge from one another, allowing us to illuminate both photodiodes.
One of the sampled beams is attenuated by an additional factor of $20\times$ by reflection off of an uncoated fused silica glass plate, such that the dynamic range of its photodiode covers $0-20$ W of the beam directed at the atoms, and controls the Berek compensator based, high-power feedback loop.
The other beam is not attenuated, save by the beam sampler, and its dynamic range covers $0-1$ W.
To prevent damage to the low power photodiode when the system is at its highest powers, a shutter (Stanford Research Systems SR475) blocks the light when the high power loop is in use.
The low power photodiode passes its signal $V_{\mathrm{pd}}$ to a loop filter circuit, which compares it to the control system's set-point, generated by digital to analog converter.
The loop filter's output controls the IF port of a microwave mixer (Minicircuits ZFM-2-S+), effectively modulating the RF power supplied by a low noise phase-locked loop (PLL).
The resulting RF signal is then amplified and sent to the AOM.
Let us now consider every part of this signal chain individually.

\subsubsection{Phase-locked loop (PLL)}
The local oscillator (LO) RF source is a custom-built printed circuit board (PCB) implementing a phase locked loop (PLL) based on the ADF4002 IC and Crystek's CVCO55CL-0060-0110 voltage controlled oscillator (VCO), preamplified with a low noise amplifier (Minicircuits PSA4-5043+).
To ensure frequency stability (which translates directly into beam pointing of the laser), the PLL is locked to the $10$ MHz clock distributed around the lab.
In addition to PLL functionality, the board also contains two digitally-controlled, high-isolation RF switches, giving a total isolation exceeding $90$ dB.

>>>>

\newcommand\unit[2]{\ensuremath{#1~\mathrm{{#2}}}}
\newcommand\FIXME{{\color{red}\ensuremath{\mathrm{FIXME}}}}

@@

\author{A. Mazurenko}
\altaffiliation{
  Present address: \FIXME{}}
\author{S. Blatt}
\altaffiliation{
  Present address: Max-Planck-Institut f{\"u}r Quantenoptik,
  Hans-Kopfermann-Stra{\ss}e 1, 85748 Garching, Germany}
\author{F. Huber}
\altaffiliation{
  Present address: IPG Photonics Corporation, Oxford, MA 01540}
\author{M. F. Parsons}
\altaffiliation{
  Present address: \FIXME{} TOO}
\author{C. S. Chiu}
\author{G. Ji}
\author{D. Greif}
\author{M. Greiner}
\affiliation{
  Department of Physics, Harvard University,
  Cambridge, Massachusetts, 02138, USA}

\date{\today}
\begin{abstract}
  We describe the design and implementation of a stable high-power \unit{1064}{nm} laser system for use in experiments with ultracold atoms trapped in optical lattices. The system is based on a low-noise laser amplified by an array of four heavily modified, high-power fiber amplifiers. The beam intensity is stabilized and controlled with a nonlinear feedback loop. Using real-time monitoring of the resulting optical lattice, we demonstrate a shot-to-shot stability of the lattice position at 1\% of the lattice spacing and a long-term (six month) relative stability of 0.5\% RMS of the lattice spacing.
\end{abstract}
\maketitle

Optical lattices holding arrays of ultracold atoms have become a powerful experimental platform for quantum simulations~\cite{Friedenauer2008, Kim2010, Struck2011, Simon2011, Yan2013, Drewes2016, Murmann2015}.
The recent advent of quantum gas microscopy~\cite{Bakr2009, Sherson2010, Haller2015, Cheuk2015, Parsons2015, Edge2015, Omran2015, Greif2016, Cheuk2016, Parsons2016, Boll2016, Cheuk2016a, Brown2017} has enabled studies with unprecedented control over individual atoms in many-body systems governed by the Hubbard Hamiltonian~\cite{Hubbard1963}, resulting in extremely low energy scales~\cite{Mazurenko2017}.
However, this progress has placed ever more stringent technical constraints on the lasers used to trap and manipulate the atomic systems.
Even if the systems thermalize, heating rates can limit the interrogation time available to probe low-temperature states~\cite{Blatt2015}.
In particular, experiments that rely on a small number of atoms held by multiple traps~\cite{Preiss2015, Choi2016, Mazurenko2017} place stringent requirements on the relative positional stability of the traps, to ensure that the operations performed by the traps have adequate fidelity.

In this work we discuss major improvements to our high power optical lattice system~\cite{Blatt2015} and show levels of intensity noise below \unit{-115}{dBc/Hz} between \unit{1}{kHz} and \unit{3}{MHz} Fourier frequency.
In addition to the intensity stability, we demonstrate a short-term (shot-to-shot) stability of the lattice position at 1\% of the optical lattice spacing, and a long-term (six month) relative stability of 0.5\% RMS of the lattice spacing itself.

@@

The architecture of the system, shown in Fig.~\ref{fig:high_level}, is based on a single low-power, low-noise seed laser that supplies light to an array of four high-power fiber amplifiers (one for a transport beam~\cite{Huber2014}, and one each for the $X,Y,Z$ lattices~\cite{Blatt2015,Parsons2016}).
These amplifiers generate enough light to perform site resolved imaging~\cite{Parsons2015}, but also substantially increase the level of intensity and pointing noise.
The fiber amplifier output is controlled by two feedback loops, which stabilize the light in low power and high power regimes, respectively, as discussed later.
After intensity stabilization, the laser beam is mode-shaped to a desired waist.
The optical system is engineered to be passively stable in terms of pointing, but the position and phase of the optical lattice are monitored in real time, and beam pointing can be controlled with remotely actuated mirrors.

@@

\section{Laser system}
\subsection{Seed laser system}
The lattice laser system is seeded by an Innolight Mephisto laser, which provides approximately \unit{2}{W} of \unit{1064}{nm} light.
This commercial laser uses a non-planar ring oscillator (NPRO) design that ensures superb phase and intensity stability.
This source is split and coupled into four single-mode, polarization-maintaining fibers (Thorlabs PM980-XP), each connected to the input of a fiber amplifier.
Three of these fiber amplifiers are used to produce the 3D optical lattice~\cite{Blatt2015,Parsons2016}, and a fourth is used for transport~\cite{Huber2014}.
To avoid interference of the lattice beams, two of the beam paths pass through acousto-optic modulators (AOMs), which detune the lattice frequencies from each other.

\begin{figure*}
  \begin{center}
    \includegraphics{Figure2.pdf}
    \caption{\textbf{Laser noise of the lattice system.} (a) Shows the laser noise of unmodified and free-running Nufern fiber amplifiers and the Mephisto seed, as well as an independent Mephisto MOPA 55W (not used for the optical lattice presented in this manuscript). (b) Shows the open-loop and closed-loop behavior of the high-power feedback loop using the  modified fiber amplifiers, as described in the main text. (c) Shows the open-loop behavior of the low-power feedback loop using the modified fiber amplifiers, including an additional low-pass filtering stage to limit the loop bandwidth. (d) Closed-loop behavior of the low-power feedback loop without additional low-pass filtering, measured on the same setup as panel (c).}\label{fig:noises}
  \end{center}
\end{figure*}

\subsection{Fiber amplifier system}
Quantum gas microscopy of Li$^6$ requires single site trap frequencies of approximately \unit{1}{MHz}~\cite{Parsons2015} to be in the Lamb-Dicke regime necessary for Raman sideband cooling.
Such trap frequencies may only be attained using tightly focused, high-power beams.
In practical terms, this requires approximately \unit{30}{W} per lattice axis, which we attain using Yb-doped fiber amplifiers (Nufern NuAmp SUB1174-34).
These amplifiers can amplify \unit{150}{mW} of seed light to up to approximately \unit{45}{W}. They use a length of doped fiber pumped with \unit{808}{nm} light from fiber coupled diode sources to amplify the \unit{1064}{nm} seed light.
The amplification is performed using a two-stage design, where the stages are controlled and powered using separate power supplies and electronics.

\begin{figure*}[t]
  \begin{center}
    \includegraphics{Figure3.pdf}
    \caption{\textbf{Optical layout} of the optical lattice system for beam shaping and modulation.}\label{fig:optical_layout}
  \end{center}
\end{figure*}

@@

Further sources of technical noise arise from the design of the amplifier. There are two power supplies in the fiber amplifier: (1) a general purpose power supply for the control electronics and the first, low-power amplification stage and (2) a high-power supply for the second, high-power stage.
Despite the presence of a water-cooled cold-plate for the gain fiber and pump diodes, the high power supply is air cooled by an internal fan with variable frequency.
This fan adds noise at acoustic frequencies and degrades pointing stability by increasing vibrations on the output fiber tip.
To reduce acoustic coupling and reduce switching power supply noise, both power supplies were removed from the fiber amplifier chassis, and the low-power supply was replaced by a linear power supply (Acopian A24MT350M).
Further, a high current low-pass filter (MPE DS26387) was added to the high-power supply (Lumina LDD-600-60-10-5VP/M).

These upgrades serve to suppress many of the noise spurs measured in the fiber amplifier RIN.
The effects can be seen when comparing Fig.~\ref{fig:noises}(a) and (b), particularly in the region of approximately \unit{10}{kHz} Fourier frequency.
The measurements in panels (a)-(d) were carried out on a single fiber amplifier, prior to and following the modifications. The initial and final noise spectra were verified to be consistent between amplifiers coming from the same manufacturing batch.
The figure is based on fiber amplifiers from a relatively new batch, which even in its unmodified configuration does not exhibit a large switching supply spur.
This is not the case for certain older batches, where the spur is approximately \unit{30}{dB} above the background.

The SBS noise is fundamental to the fiber amplifier architecture and cannot be entirely avoided. However, since it is a function of fiber length and optical power~\cite{Agrawal2013}, it can be minimized. To this end, we use custom-ordered amplifiers with a \unit{50}{cm} long output fiber, compared to the standard length of a few meters.
In addition, we use the lowest pump diode current that yields enough power to perform imaging.
The rise of the noise floor as a function of the pump diode current can be seen in Fig.~\ref{fig:noises}(a) at Fourier frequencies above \unit{100}{kHz}.

The operation of the Nufern fiber amplifiers can be further optimized by finely controlling the cooling water temperature.
The water temperature tunes the wavelength of the light supplied by the diodes that optically pump the gain fiber.
Optimizing the temperature is advantageous because (1) the maximum output power of the fiber amplifier is increased due to the improved efficiency of pump light conversion and (2) the lifetime of the amplifier is increased because less stray pump light needs to be absorbed when the pump light is separated from the output, reducing the thermal load.

@@

The large-diameter fiber tip at the output of the fiber amplifier is mounted in a monolithic mount made from copper.
The beam is collimated using an $f=\unit{20}{mm}$ fused-silica, air-spaced triplet collimator (Opto-Sigma HFTLSQ-15-20PF1) angled with respect to and centered on the angle-cleaved fiber tip.
The polarization is set using an IBS coated thin-film plate polarizer (Precision photonics PO1056-FY), which has the added benefit of rejecting the undesired cladding modes of the fiber, which carry about \unit{1}{W} of light.
The fact that the lattice is retro-reflected, combined with the observation that optical isolation falls with applied optical power~\cite{Yoshida1999} means that two stages of optical isolators must be used.
The optical isolators are based on \unit{5}{mm} aperture Tb$_3$Ga$_5$O$_12$ (TGG) isolators (Thorlabs IO-5-1064-VHP), where we replaced the default polarizers (BK7 Brewster plates) by IBS-coated fused silica plate polarizers.
The isolators are located approximately \unit{1}{m} away from the ultracold atoms which are susceptible to stray magnetic fields.
To overcome this problem the isolators are enclosed in mu-metal shielding.
Actuation of the power is performed after the optical isolators, such that the full fiber amplifier output continuously passes through them, minimizing thermal transients during the experimental sequence.

@@

After the beam has been cleaned up and isolated, it is important to consider two aspects of the desired experiment:
\begin{enumerate}
\item The lattice power must be continuously tunable from around \unit{10}{\mu{}W} to around \unit{20}{W}, corresponding to lattice depths in the approximate range of $10^-3$ to $2 \times 10^3 E_R$, where $E_R$ is the geometric recoil energy corresponding to the lattice spacing~\cite{Blatt2015}.
\item The location of the minimum Gaussian beam waist must be accurately positioned to overlap the other lattices and dipole traps present in the experiment, and must remain aligned for the duration of the data collection process (hours to days).
\end{enumerate}
To satisfy the first of these requirements, a two-stage feedback system is implemented.
Two stages are used because the experiment typically operates in one of two regimes, which we term \emph{detection} and \emph{interaction}.
In the interaction phase of the experiment, the lattice is relatively shallow ($\approx \unit{100}{mW}$, corresponding to depths of $\approx 10 E_R$), which allows atoms to tunnel and interact with each other.
In this regime, precise and potentially fast control is required.
In the detection phase, the depth of the lattice is dramatically raised (to $\approx 2000 E_R$), which isolates atoms in their individual wells so that Raman sideband imaging can be performed~\cite{Parsons2015}.
In the detection regime, the control need not be fast, and the passive stability of the Mephisto laser ensures that noise is low at relevant frequency scales ($\approx\unit{2}{MHz}$ Fourier frequency).
Thus, two control loops are used, a fast loop based on an acousto-optic modulator that controls the laser at low powers, and a slow loop which uses a Berek compensator to control the laser at high powers.

After the isolation stage, we use the sequence of optical elements shown in Fig.~\ref{fig:optical_layout}, consisting of a polarizer, half-wave plate (HWP), Berek compensator mounted on a precision galvo (Thorlabs GVS002, or Camtech 8320K, depending on the beam path) half-wave plate, and polarizer.
The Berek compensator is a $z$-cut quartz plate anti-reflection-coated at \unit{1064}{nm}.
By tilting the plate about its vertical axis by a few degrees, the extraordinary axis is mixed into the propagation of the beam, leading to tilt-dependent birefringence that can be tuned from a zero-wave plate past a half-wave plate, represented by a retardation angle $\phi$.
Combined with the subsequent polarization optics, rotation of the Berek compensator can vary the transmitted power with a contrast of $\sim$1000:1.
Figure~\ref{fig:berek} shows the transmitted power as a function of retardation angle $\phi$ and fast-axis angle $\theta$ of the waveplates surrounding the compensator.
\begin{figure}
  \begin{center}
    \includegraphics{Figure4.pdf}
    \caption{\textbf{Berek compensator} transmission as a function of the Berek compensator's retardation angle $\phi$ and waveplate rotation angles $\theta$.}\label{fig:berek}
  \end{center}
\end{figure}

@@

\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{Figure5.pdf}
    \caption{\textbf{Berek compensator based feedback loop}. This Berek compensator based feedback loop is used to vary the beam intensity at relatively high power. (a) Shows the small signal impulse response and (b-c) show the corresponding Bode plots of the transfer function $G$.}\label{fig:berek_step_response}
  \end{center}
\end{figure}
A key point, as seen in Fig.~\ref{fig:berek}, is that if the waveplates are rotated together, the maximum transmitted power is fixed, while the minimum can be raised.
The waveplates are detuned such that the maximum power needed in \emph{interaction} mode is the minimum power transmitted by the Berek compensator system, strongly suppressing fluctuations due to angle positioning noise of the Berek compensator.
Naturally, rotation of a parallel plate in the beam path induces a shift of the beam, but this is acceptable because the Berek compensator is only active during the imaging phase of the experiment, when the position of the underlying harmonic trap is largely irrelevant.
Fortunately, the spatial phase of the optical lattice is set by the retro-reflector mirror, which does not move, and is not affected by the rotation of the Berek compensator.
This system functions as the actuator arm of a digital feedback loop, where the power of the beam is measured using a low noise photodiode. This feedback loop regulates the lattice between \unit{1}{W} of optical power and its maximum value (which depends on the amplifier, but is $\approx$\unit{20}{W}).
The step response and corresponding Bode plots~\cite{Bechhoefer2005} are shown in Fig.~\ref{fig:berek_step_response}.
A polarization rotation by $2 \pi$ is produced by approximately $3^\circ$ of axial rotation of the quartz plate.
Since we never need to rotate the plate outside this range, the position is constrained to this window in software.

Because two feedback loops are present in the system (the second one detailed in the next section), care must be taken so that they do not function simultaneously.
To accomplish this, the low-power loop is always manually railed to its high extreme when the high-power loop is active, and the digital high-power loop is disabled when the low-power loop is active.

\subsection{Low-power feedback}
The high-power feedback used in the \emph{imaging} phase of the experiment is exceedingly robust because the heating rate from scattered resonant imaging light far exceeds any technical heating from the loop~\cite{Blatt2015}.
Even if the loop was exceedingly noisy, the heating would be entirely compensated by the simultaneous Raman cooling~\cite{Parsons2015}.

In contrast, the low-power loop that controls the lattice laser during the \emph{interaction} phase of the experiment is critically important:
\begin{enumerate}
  \item Its effect on the RIN can limit the system temperature.
  \item It can affect the pointing.
  \item It must have closed-loop bandwidths of $\geq\unit{1}{kHz}$, and open-loop modulation bandwidths $\geq \unit{100}{kHz}$.
\end{enumerate}
To this end, we use an all-custom loop based on a TeO$_2$ AOM (Gooch \& Housego AOMO 3080-198), where the power of the RF supplied to the AOM is used to actuate the fast, low-power feedback loop.
The AOM was chosen for its low thermal lensing and easily-accessible RF powers.
It is mounted on a monolithic, water-cooled flexure mount that allows optimization of the AOM efficiency, while remaining stable over years of operation.

\begin{figure}
  \begin{center}
    \includegraphics{Figure6.pdf}
    \caption{\textbf{Feedback circuits.} (a) The reverse-biased photodiode. (b) Externally referenced phase-locked loop. (c) Nonlinear loop filter. (d) RF power modulation using a mixer.}\label{fig:circuits}
  \end{center}
\end{figure}
The feedback circuit is shown in Fig.~\ref{fig:circuits}(c). As seen in Fig.~\ref{fig:optical_layout}, a pair of low-noise photodiodes [Fig.~\ref{fig:circuits}(a)] measures the optical power of the beam sampled with a weakly reflective optic (Thorlabs BSF10-C).
The beam sampler is wedged in such a way that the two reflections diverge from one another, allowing us to illuminate both photodiodes.
One of the sampled beams is attenuated by an additional factor of 20 by reflection off of an uncoated fused silica glass plate, such that the dynamic range of its photodiode covers \unit{0-20}{W} of the beam directed at the atoms, and controls the Berek-compensator-based, high-power feedback loop.
The other beam is not attenuated, save by the beam sampler, and its dynamic range covers \unit{0-1}{W}.
To prevent damage to the low-power photodiode when the system is at its highest powers, a shutter (Stanford Research Systems SR475) blocks the light when the high-power loop is in use.
The low-power photodiode passes its signal $V_{\mathrm{pd}}$ to a loop filter circuit [Fig.~\ref{fig:circuits}(c)], which compares it to the control system's set-point, generated by a digital-to-analog converter.
The loop filter's output controls the IF port of a microwave mixer [Minicircuits ZFM-2-S+, see Fig.~\ref{fig:circuits}(d)], to modulate the RF power supplied by a low-noise phase-locked loop [PLL, see Fig.~\ref{fig:circuits}(b)].
The resulting RF signal is then amplified and sent to the AOM.
Let us now consider every part of this signal chain individually.

\subsubsection{Phase-locked loop (PLL)}
The local oscillator (LO) RF source is a custom-built printed circuit board (PCB), schematically shown in Fig.~\ref{fig:circuits}(b), that implements a phase locked loop (PLL) based on the ADF4002 IC and a low-phase-noise voltage-controlled oscillator (VCO, Crystek CVCO55CL-0060-0110), preamplified with a low-noise amplifier (Minicircuits PSA4-5043+).
To ensure frequency stability (which translates directly into beam pointing of the laser), the PLL is locked to the \unit{10}{MHz} frequency reference distributed around the lab.
In addition to PLL functionality, the board also contains two digitally-controlled, high-isolation RF switches, giving a total isolation exceeding \unit{90}{dB}.
